---
title       : Gestational age calculator
subtitle    : A guide for the shiny app
author      : Juan M Zambrano
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Introduction

This [simple yet useful calculator](https://jmzam.shinyapps.io/gestationalcalculator) can estimate gestational age and estimated date of delivery based on a current (or otherwise desired) date and one of two parameters:

1. Date of first day of last menstrual period
2. Date of ultrasound and gestational age by ultrasound*

The gestational age is calculated using the calendar method (i.e. counting how many days have passed since the last menstrual period or the ultrasound date). 

The estimated date of delivery is calculated by adding 280 days to the last menstrual period [1].

A more detailed explanation of why these methods are useful for you and your doctor can be found [here](http://rhrealitycheck.org/article/2013/10/17/whats-in-a-week-pregnancy-dating-standards-and-what-they-mean/). More details on how ultrasound can be used to calculate gestational age can be found [here](https://tonygood4.wordpress.com/2013/02/03/developing-and-eye-for-ultrasound/).



--- .class #id 

## Estimated delivery date by last menstrual period

As an example, let us consider a last menstrual period 80 days ago, the calculation would be:

```{r eddlmp}
 lmp <- Sys.Date()-80 #Example last menstrual period 80 days ago
 #Estimated date of delivery = last menstrual period + 280
 edd <- paste("The estimated date of delivery is", lmp+280)
 # Calculates time difference between today and last menstrual period
 diff <- as.numeric(Sys.Date()-lmp) 
 # Current gestational age in weeks and days
 cgalmp <- paste("The current gestational age is",diff%/%7, "weeks", diff%%7, "days") 
 cgalmp
 edd
 
```


--- .class #id 

## Estimated delivery date by ultrasound

But what if we cannot trust the last menstrual period (because it cannot be remembered or cycles are irregular). If we had an ultrasound done 3 weeks ago, stating that the estimated gestational age is 7 weeks 5 days, suggesting a different gestational age and date of delivery. The estimated date of delivery and current gestational age would be calculated as follows:


```{r}
        today <- Sys.Date(); usdate <- today - 21; usweek <- 7 ; usday <- 5
        diff <- as.numeric(as.Date(today)-as.Date(usdate))+7*usweek+usday
        cgaus <- paste("The current gestational age is", diff%/%7, "weeks", diff%%7, "days")
        eddus <-paste("The estimated date of delivery is", as.Date(today)+(280-diff))
        cgaus
        eddus 
```


--- .class #id 

## References


1. World Health Organization. ICD-10: International statistical classification of diseases and related health problems, 10th revision. Volume 2. 2nd ed. Geneva: WHO; 2004. Available at: http://www.who.int/classifications/icd/ICD-10_2nd_ed_volume2.pdf. Retrieved November 12, 2015.

Thank you for viewing the presentation. Be sure to visit the [app! :)] (https://jmzam.shinyapps.io/gestationalcalculator)
